---
title: "hw2"
author: "SamG"
date: "March 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(babynames)
```
Q1
```{r}
pop_names <- babynames %>%
  group_by(name) %>%
  summarize(x = sum(n)) %>%
  arrange(-x)
pop_names[1:10,]
```

Q2
After a lot of the major wars (WWI, WW2) there were spikes in the birth rate and during periods of economic growth
```{r}
babynames %>%
  group_by(year) %>%
  summarize(births = sum(n)) %>%
  ggplot(aes(x = year,
         y = births)) +
  geom_col()
```

q3
A was the most common
```{r}
A_names <- babynames %>%
  mutate(a = (str_detect(babynames$name, "a|A"))) %>%
  mutate(a1 = ifelse(a == TRUE, 
         1,
         0)) %>%
  summarize(sumA = sum(a1))

E_names <- babynames %>%
  mutate(E = (str_detect(babynames$name, "e|E"))) %>%
  mutate(e1 = ifelse(E == TRUE, 
                     1,
                     0)) %>%
  summarize(sume = sum(e1))

I_names <- babynames %>%
  mutate(I = (str_detect(babynames$name, "i|I"))) %>%
  mutate(i1 = ifelse(I == TRUE, 
                     1,
                     0)) %>%
  summarize(sumi = sum(i1))

O_names <- babynames %>%
  mutate(o = (str_detect(babynames$name, "o|O"))) %>%
  mutate(o1 = ifelse(o == TRUE, 
                     1,
                     0)) %>%
  summarize(sumo = sum(o1))

U_names <- babynames %>%
  mutate(u = (str_detect(babynames$name, "u|U"))) %>%
  mutate(u1 = ifelse(u == TRUE, 
                     1,
                     0)) %>%
  summarize(sumu = sum(u1))
A_names[[1,1]]
data = c(A_names[[1,1]], E_names[[1,1]], I_names[[1,1]], O_names[[1,1]], U_names[[1,1]])
df <- data.frame("vowel" = c("A", "E", "I", "O", "U"), "count" = data)
df %>%
  ggplot(aes(x = vowel,
             y = count)) +
  geom_col()
```


Q4
Maryelizabeth and Michaelangelo
```{r}
babynames %>%
  group_by(name) %>%
  summarize(number_of_names = sum(n)) %>%
  filter(number_of_names >1000) %>%
  mutate(name_length = str_length(name)) %>%
  arrange(-name_length)
```


Q5
Today baby names are shrinking to return to "normal", as a response to the spike in the 1900's. The spike could have been due to societal trends, and people could have thought that longer names were more formal and pretty sounding at the time. 
```{r}
babynames %>%
  mutate(length = str_length(babynames$name)) %>%
  group_by(year) %>%
  summarize(total_people = sum(n), total_lengths = sum(length * n)) %>%
  mutate(avg.length = total_lengths / total_people) %>%
  ggplot(aes(x = year,
             y = avg.length)) +
  geom_line()
```

q6
```{r}

babynames %>%
  mutate(begins_with_vowel = str_detect(name, "^A|^E|^I|^O|^U")) %>%
  mutate(begins.vowel = ifelse(begins_with_vowel == TRUE,
                               n,
                               0)
  ) %>%
  mutate(ends_with_vowel = str_detect(name, "a$|e$|i$|o$|u$")) %>%
  mutate(ends.vowel = ifelse(ends_with_vowel == TRUE,
                               n,
                               0)) %>%
  group_by(year) %>%
  summarize(births = sum(n), vowel.start = sum(begins.vowel), vowel.end = sum(ends.vowel)) %>%
  mutate(proportion.with.vowelstart = vowel.start / births, proportion.with.vowelend = vowel.end / births)%>%
  gather(proportion.with.vowelstart, proportion.with.vowelend, key = vowelLoc, value = proportion) %>%
  ggplot(aes(x = year,
             y = proportion)) +
  geom_line(aes(color = vowelLoc))
```

